---
title: "Describing a Wine"
author: "Ellis Hughes"
date: "June 14, 2018"
output:
  html_document:
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

After the grape type and winery, the descriptions are the huge part of the wine purchasing process. Usually these descriptions draw from the ["Wine Aroma Wheel"](https://wcicdn-thewinecellarins.netdna-ssl.com/wp-content/uploads/2010/05/Davis-Wine-Aroma-Wheel1.jpg) developed at UC Davis. It categorizes the scents wines can have, from the pleasant to the less so.

In this analysis I will be using the "tidytext" packge developed by David Robinson and Julia Silge to see if I can find insight in how wines are described across varietals. As in my prior analysis of wine, the data used for this analysis can be found in csv generated by [Zack Thoutt]("https://www.kaggle.com/zynicide/wine-reviews/data").

```{r load_data}
library(knitr)
library(tidyverse)
library(tidytext)
library(tokenizers)
opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
wineData<-read.csv("../winemag-data-130k-v2.csv",header = TRUE,stringsAsFactors = FALSE)
```

#The Breakdown

After reading in the csv, I performed some basic filtering to make sure we didnt have any countries or varietals that were too rare. each country needed a minumum of 500 representatives, and each variatal 300. This left me with 113601 observations.

Next, the unnest_tokens was used to split the descriptions into individual words and any stop words were removed. 

```{r extract_descriptions}
wineData_filtered<-wineData%>%
  group_by(country)%>%
  filter(n()>500)%>%
  ungroup()%>%
  group_by(variety)%>%
  filter(n()>300)%>%
  ungroup()%>%
  select(-X,-taster_twitter_handle)

wineData_TidyDescriptions<-wineData_filtered%>%
  unnest_tokens(word, description)%>%
    anti_join(get_stopwords())
  
```

The top counted words remaining make sense, as the word "wine", when describing wine, is logical. And since the desciption is for describing the taste and smell of the wine, much more refined words such as "flavors" and "aromas" are used. Fruit is a common word in describing wine, as it is a desireable quality, so it finds itself as the third most commonly used word in descriptions. 

```{r description_counts}

wineData_TidyDescriptions %>%
  count(word, sort = TRUE) 

```

# Description By Varietal

I removed some of the generic words from the set, which showed up commonly in the descriptions in general such as "wine", in an effort to find the actual descriptors of the wine. 

Looking at the results, the words that I think of when I think of a varietal are common descriptors. Chardonnay was described with words like fruit, apple, oak and  Merlot as cherry, tannins, black, and finish. If this is what the wine makers are going for, they succeeded! 

```{r description_counts_varietal}
wineData_TidyDescriptions %>%
  filter(!word%in%c("wine","flavors","aromas","drink"))%>%
  group_by(variety)%>%
  count(word,sort=TRUE)%>%
  {
    uniqueWines<-unique(.$variety)
    top5wordsMatrix<-data.frame(matrix(NA,ncol=5,nrow=length(uniqueWines)),row.names = uniqueWines)
    colnames(top5wordsMatrix)<-c("Word1","Word2","Word3","Word4","Word5")
    for(var in uniqueWines){
      tmp<-.[.$variety==var,"word"]%>%filter(!toupper(word)%in%toupper(strsplit(var," ")[[1]]))%>%unlist
      top5wordsMatrix[var,]<-tmp[1:5]
    }
    return(top5wordsMatrix)
  }%>%
  mutate(wine=rownames(.))%>%
  arrange(wine)%>%
  select(wine,Word1,Word2,Word3,Word4,Word5)

```

# Sentiment vs Score

I was now curious if the writers seniment would change when you give a wine a higher or lower score. Overall, surprisingly once the wine reached a certain point, the reviewers could only be so positive. Unless, of course, the wine was perfect. In whichcase, the reviewer had very little negative comments. 

We must also remember that sentiment analysis did no take into account conditioning words such as not or no. Also, some flavor descriptors are found to be negative in sentiment that is actually positve, such as "lemon". 

```{r sentiment vs score}

bing <- get_sentiments("bing")

wine_sentiments<-wineData_TidyDescriptions %>%
  inner_join(bing) %>%
  filter(!word%in%c("wine","flavors","aromas","drink"))%>%
  mutate(points=factor(as.character(points),levels = unique(as.numeric(as.character(points)))),
         sentiment_positive=ifelse(sentiment=="positive",1,-1))%>%
  group_by(points)%>%
  summarise(sentiment_positive=mean(sentiment_positive))%>%
  mutate(points=as.numeric(as.character(points)))%>%
  arrange(points)

ggplot(wine_sentiments,aes(x=points,y=sentiment_positive))+
  geom_point(aes(color=I("blue")))+
  geom_hline(yintercept = 0)+
  geom_smooth()+
  geom_text(aes(x=95,y=.025,label="Neutral Sentiment"))+
  ggtitle("Average Sentiment by Wine Score")

```


#Obligatory Word Cloud!

Le random wordcloud, colored the words based on which word is most likely to be related to a white or red wine! I'll let you decide which color represents which.

```{r wordCloud}

library(wordcloud2)

whites<-c("Albariño","Bordeaux-style White Blend","Champagne Blend","Chardonnay","Chenin Blanc","Gewürztraminer","Glera","Grüner Veltliner","Moscato","Pinot Blanc","Pinot Grigio","Pinot Gris","Portuguese White","Rhône-style White Blend","Riesling","Sauvignon Blanc","Sparkling Blend","Viognier","White Blend")
reds<-c("Aglianico","Barbera","Bordeaux-style Red Blend","Cabernet Franc","Cabernet Sauvignon","Carmenère","Corvina, Rondinella, Molinara","Gamay","Garnacha","Grenache","Malbec","Merlot","Nebbiolo","Nero d'Avola","Petite Sirah","Pinot Noir","Port","Portuguese Red","Red Blend","Rhône-style Red Blend","Sangiovese","Sangiovese Grosso","Sauvignon","Shiraz","Syrah","Tempranillo","Tempranillo Blend","Zinfandel")
rose<-c("Rosé")

wine_COUNTS<-wineData_TidyDescriptions %>%
  mutate(variety=ifelse(variety%in%whites,"white",
                 ifelse(variety%in%reds,"red",
                 ifelse(variety%in%rose,"rose",variety))))%>%
  filter(!tolower(word)%in%c("wine","flavors","aromas","drink",tolower(unique(variety))),tolower(variety)%in%c("red","white"))%>%
  group_by(word)%>%
  summarise(REDS=sum(variety=="red"),
            WHITE=sum(variety=="white"),
            ROSE=sum(variety=="rose"),
            COUNT=n()
            )%>%
  data.frame()%>%
  mutate(REDS=REDS/sum(wineData_filtered$variety%in%reds),
         WHITE=WHITE/sum(wineData_filtered$variety%in%whites),
         ROSE=ROSE/sum(wineData_filtered$variety%in%rose))%>%
  arrange(COUNT)%>%
  mutate(color=ifelse(REDS>WHITE,"#59143b","#f8fce0"))


wine_COUNTS%>%
  arrange(desc(as.numeric(COUNT)))%>%
  mutate(freq=COUNT)%>%
  top_n(300)%>%
  wordcloud2(color = wine_COUNTS$color,backgroundColor = "black",fontFamily = "Bookman walter")


```